---
title: "Time Series Data Visualization"
---

```{r}
#| code-fold: true
#| code-summary: 'Importing Libraries'
#| warning: false
#| output: false
library(tidyverse)
library(ggplot2)
library(forecast)
library(astsa) 
library(xts)
library(tseries)
library(fpp2)
library(fma)
library(lubridate)
library(tidyverse)
library(TSstudio)
library(quantmod)
library(tidyquant)
library(plotly)
library(ggplot2)
library(imputeTS)
library(gridExtra)
library(reticulate)
library(readxl)
```

# Time Series Data Visualization

```{r, warning=FALSE, message=FALSE, echo=FALSE}
# Read the data from the CSV file
new_yearly <- read_csv("new_yearly.csv",show_col_types = FALSE)
```



```{r, warning=FALSE, message=FALSE}
#| code-fold: true
#| code-summary: 'Average Nitrate Levels'
#| warning: false
#| output: true

# Create the ggplot
p <- ggplot(new_yearly, aes(x = year, y = mean)) +
  geom_point(color = 'black', size = 1) +  # Add the data points
  geom_line(color = 'dodgerblue2', linewidth = 1) +   # Add the line with updated syntax
  geom_smooth(method = 'lm', color = 'red', linetype = "dashed") + # Add a smooth line
  scale_x_continuous(breaks = new_yearly$year) +  # Ensure all years are shown
  theme_minimal() +  # Use a minimal theme
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    plot.title = element_text(hjust = 0.5)  # Center align the title
  ) +
  labs(
    title = 'Average Nitrate Levels in the Drinking Water in California',
    x = 'Year',
    y = 'Average Nitrate [mg/L]'
  )

# Convert ggplot to a plotly object for interactivity
p_plotly <- ggplotly(p)

p_plotly
```


The yearly graph above shows the average nitrate levels in California's drinking water from 1998 to 2011. The points represent annual average nitrate concentrations, the red dashed line, representing a linear regression model, has been added to indicate the overall trend within the observed period. 

This helps us identify potential patterns, trends, or anomalies in nitrate levels over the years. Notably, it's crucial to observe how these levels fluctuate and consider external factors that might influence these changes, such as environmental regulations or significant events affecting water quality.



# Tableau Dashboard

#### Interactive County-level Nitrate Level Map

The Tableau dashboard below is an interactive exploration of nitrate levels across various counties in California. This map helps in providing insights into regional disparities or areas of particular concern. There is time slider feature where you can see how nitrate levels in each county have evolved from 1998 to 2011.


<iframe src="https://public.tableau.com/views/california_17104055601850/Dashboard1?:embed=y&:display_count=yes&:showVizHome=no" width="100%" height="1200"></iframe>



